Safe and Robust Formal Synthesis of Digital Controllers for Continuous Plants with Transient Performance Specifications

### 2017-06-21 automatica/paper.tex
Formatting changes

\section{Introduction}
Modern implementations of embedded control systems have proliferated with
the availability of low-cost devices that can perform highly non-trivial
control tasks, with significant impact in numerous application areas such as
life sciences, environmental control and robotics~\cite{astrom1997computer, Franklin15}. 
Correct synthesis is non-trivial, even in cases with restriceted dynamics.
We examine the case of Linear Time Invariant (LTI) Models, for which the
synthesis of controllers is well understood. However, the
use of digital control architectures adds new challenges caused by artifacts
specific to digital control, such as the effects of finite-precision arithmetic, time
discretization and quantization noise introduced by A/D and D/A conversion.

@@

\item We automatically generate {\em correct-by-construction} digital
  controllers using an inductive synthesis approach.  Our application of program
  synthesis is non-trivial and addresses challenges specific to control
  systems, such as quantizers and FWL.  In particular, we
  have found that a two-stage verification engine that continuously refines
  the precision of the fixed-point representation of the plant yields a
  speed-up of two orders of magnitude over a conventional one-stage
  verification engine.

%
\item  We compute state-feedback controllers that 
  guarantee a given safety property alongside stability and response time
  specifications.  Existing methods for controller synthesis rely on transfer
  function representations, which are inadequate to prove safety requirements.
%
\item We provide three novel algorithms: the first, na\"ive one, relies on
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction refinement and
  acceleration to improve scalability while retaining soundness.  A third algorithm
  evaluates systems using transfer functions where no state-space description is given.
  All three approaches provide sound synthesis of feedback controllers and consider
  the various sources of imprecision in the implementation of the control
  algorithm and in the modeling of the plant.
%
\item We develop a model for different sources of quantization error and
  their effect on reachability properties.  We give bounds that ensure the
  safety of our controllers in a hybrid continuous-digital domain.
\end{enumerate}

@@

%-------------------------------
\subsection{Abstract Acceleration} 
\label{ssec:aa}
%-------------------------------
For abstraction based synthesis, we will use Abstract Acceleration~\cite{JSS14,cattaruzza2015unbounded}.
Given an iterative program with dynamics
\begin{equation}
\vec{x}_{k+1}=\mat{A}\vec{x}_k : \vec{x}_k \in \mathbb{R}^n \wedge \mat{A} \in \mathbb{R}^{n \times n} \wedge k \in [0\ \ \infty),
\end{equation}
we want to find all possible states visited over an infinite time horizon starting at an initial set $X_0$. We call this set the \emph{reach tube} of the model:
\begin{equation}
\hat{X} = \{ \vec{x}_k : k \in [0\ \infty) \wedge \vec{x}_0 \in X_0 \wedge \vec{x}_{k+1}=\mat{A}\vec{x}_k\}.
\end{equation}
It is easy to see that the infinite number of iterations would result in an unbounded cumulative rounding error, but using Abstract Acceleration we transform this problem into a one-step solution.
We first change the problem statement by means of acceleration into 
\begin{equation}
\hat{X} = \{ \vec{x}_k=\mat{A}^k\vec{x}_0 : k \in [0\ \infty) \wedge \vec{x}_0 \in X_0 \}, 
\end{equation}
and further define new model semantics and corresponding set: 
\begin{equation}\label{eq:aa_reachtube}
\hat{X} \subseteq \hat{X}^\sharp = \mathcal{A}X_0, \text{ such that } \bigcup_{k \in [0\ \infty)} \mat{A}^k \subseteq \mathcal{A}, 
\end{equation}
where the newly introduced \emph{abstract reach tube} $\hat{X}^\sharp$ is an over-approximation of the reach tube of the model \cite{JSS14}.
All we need to do is define the nature of $\mathcal{A}X_0$.

@@

Introduce set $X=\{ \vec{x} : \mat{C}\vec{x} \leq \vec{d}\}$. 
The linear transformation $\vec{x}'=\mat{S}^{-1}\vec{x}$ leads to set 
\begin{equation}
X'= \mat{S}^{-1}X = \{\vec{x}' : \mat{C}\mat{S}\vec{x}'\leq \vec{d}\}.  
\label{eq:eigenInit}
\end{equation}
Analogously, 
let $X_0'=\mat{S}^{-1}X_0$ and ${\hat{X}'^\sharp}=\mat{S}^{-1}{\hat{X}^\sharp}$ be the initial set and the abstract reach tube mapped onto the eigenspace of matrix $\mat{A}$. 
We may transform equation \eqref{eq:aa_reachtube} into:
\begin{equation}
\hat{X}' \subseteq {\hat{X}'^\sharp} = \mathcal{J}X_0', \text{ such that } \bigcup_{k \in [0\ \ \infty)}\mat{J}^k \subseteq \mathcal{J}, 
\end{equation}
where $\hat{X}'=\mat{S}^{-1}\hat{X}$ and $\mathcal{J}$ is a convex polyhedron representing restrictions on the eigenvalues of $\mat{A}$.
$\hat{X}'^\sharp$ is calculated by applying a simplex algorithm with a tableau defined by $\mathcal{J}$ and objective functions derived from each vertex of $X_0'$ (step \eqref{alg:getReachTube} in Algorithm \ref{alg:findReachTube} below).
We note, however, that if we wish to use numerical analysis in the decomposition of $\mat{A}$, 
the Tableau in $\mathcal{J}$ contains small intervals given by the errors, which cannot be processed efficiently by a regular simplex and require a new procedure.

The calculation of the abstract reach tube via abstract acceleration is encompassed in Algorithm \ref{alg:findReachTube}, 
which can be summarised as follows (its steps are denoted in parentheses):
%
\begin{enumerate}
\item We perform unsound eigen-decomposition using an existing algebra package (lines \ref{alg:calcEigenvalues}-\ref{alg:calcEigenvectors}).
\item Then we restore soundness to the results using the methods described in Section \ref{sec:eigensolver} (lines \ref{alg:soundEigenvalues}-\ref{alg:soundEigenvectors}). 
\item The inverse of the matrix of eigenvectors is calculated after soundness is restored, using interval arithmetic in order to ensure its soundness (line \ref{alg:inverse}).
\item The abstract dynamics are obtained by evaluating the convex hull of all powers of eigenvalues up to the desired number of iterations as described in \cite{cattaruzza2015unbounded} (line \ref{alg:aa}).
\item Using Equation \eqref{eq:eigenInit}, transform the initial state into the eigenspace (line \ref{alg:eigeninit}).
\item Extract the vertices of the eigen-polyhedron $X_0'$ using the sound over-approximation of the double description algorithm \cite{fukuda1996double} from Section \ref{sec:dd} (line \ref{alg:getVertices}). It is worth noting that this algorithm uses a simplex to seed it: we will first discuss the simplex algorithm, which is also needed at the next step, and then the vertex enumeration algorithm used at this stage. 
\item Calculate the mapped \emph{abstract reach tube} ${\hat{X}'^\sharp}$ that over-approximates the image of the reach tube in the eigenspace (line \ref{alg:getReachTube}). This is achieved by evaluating a set of objective functions using the abstract dynamics as a simplex Tableau, 
and via a sound simplex described in Section \ref{sec:simplex}. 
The objective functions are defined by the vertices of $X'_0$, 
denote as $V_0$, 
and by the desired template directions, 
such that 
\begin{equation*}
\vec{w}_{ij}= \vec{v}_i \circ \vec{t}_j : \vec{v}_i \in V_0 \wedge \vec{t}_j \in T,
\end{equation*}
where $T$ is the set of template directions, and $\circ$ denotes a component-wise multiplication yielding a vector. 
\item Using Equation \eqref{eq:eigenInit}, find the reach tube $\hat{X}^\sharp=\mat{S}{\hat{X}'^\sharp}$ (line \ref{alg:getNormReachTube}).
\end{enumerate}

@@

%-------------------------------
 \subsection{Overapproximating continuous time dynamics using discrete models}
 \label{sec:cont_aa}
%-------------------------------
Thus far we wave discussed the Abstract Accelertion of Discrete Time systems. However, our plant models derive
from the discretization of continuous time systems. We therefore seek to establish a sound
overapproximation which does not only encompass the selected discretization, but any chosen discretization in the 
time domain.

A dynamical system is a system in which a function describes the progression of a state over time. 
In a continuous domain with linear dynamics, it is described by a first order Ordinary Differential Equation (ODE).
\begin{equation}
\dot{x}(t)=\mat{A}_o\vec{x}(t)+\mat{B}_o\vec{u}(t) +\vec{w}(t)
\label{eq:dynamical}
\end{equation}
where $\vec{w}$  is a continuous zero-mean white noise source with covariance $N ( 0 , Q )$
Furthermore, a control system may have a derived output that is a linear combination of its states and inputs, 
which may restricts the observability of the statespace from the output space.
\begin{equation}
\vec{y}(t)=\mat{C}_o\vec{x}(t)+\mat{D}_o\vec{u}(t)
\end{equation}
Discretization of a continuous dynamical system turns the ODE into a difference equation, assuming zero-order hold
for the input $\vec{u}$ and continuous integration for the noise $\vec{w}$, to
\begin{align}
\label{eq:discretization}
\vec{x}_{k+1} &= \mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k + \vec{w}_k\\
y_k &= \mat{C}_d \vec{x}_ k + \mat{D}_d \vec{u}_ k 
\end{align}
with covariance for $\vec{w}_k$, $N ( 0 , Q_d )$
where
\begin{align}
\label{eq:discretize}
\mat{A}_d &= e^{\mat{A}_o T_s} = \mathcal{L}^{-1} { ( s \mat{I} - \mat{A}_o )^{-1} }_{t = T_s}\\
\mat{B}_d &= \int_{t = 0}^{T_s} e^{\mat{A}_o t} dt\ \mat{B}_o = \mat{A}_o^{-1} ( \mat{A}_d - \mat{I} ) \mat{B}_o\\
\mat{C}_d &= \mat{C}_o\\
\mat{D}_d &= \mat{D}_o \\
Q_d &= \int_{t = 0}^{T_s} e^{\mat{A}_o t} Q e^{\mat{A}_o^T t} dt
\end{align}
and $T_s$ is the sample time. Then
\begin{align*}
x(kT)=x_k \wedge y(kT) = y_k, \forall k
\end{align*}

@@

 Let $\mat{J}$ be a diagonal matrix, such that 
 $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
 Let us now take a different sample rate $T_2$, such that 
 $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
 We can then say that 
 \begin{equation}
 {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
 \end{equation}
 This proves the theorem for $gm(\lambda_i)=1$.
 We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
 referenced to the original eigenvalues modified by the sampling rate.
  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix. Then
\begin{align}
 \mat{J}_1=&\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k\nonumber\\
 =&\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
 \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
\vdots & \ddots & \ddots & \vdots \\
& &  &\lambda_i^k \\
\end{array} \right ] T_1^k
\end{align}
The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
\begin{align}
\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
\end{align}
which completes the proof for $gm(\lambda_i)>1$.
\end{proof}

@@

In the case of Discrete dynamics, support functions can be found by evaluating hyperplanes of consecutive exponents of the eigenvalues using differences.
In the case of continuous time, this translates to calculating derivatives. Once again, when the sets are known to be convex,
we may calculate the normal vector to this derivative and evaluate its support function on the selected iteration. The direction of the vector is determined by evaluating the support function of another point in the curve, which is known to be smaller than that of the selected point.
The corresponding derivatives are:
\begin{enumerate}
\item for real eigenvalues: we first calculate the slope of the tangent between the two progressions: $\frac{d f(k)}{d g(k)}=\frac{f'(k)}{g'(k)}$. Hence $\frac{d \lambda_1^k}{d \lambda_2^k}=\frac{log(\lambda_1) \lambda_1^k}{log(\lambda_2) \lambda_2^k}$. From this we know that the polar angle of the support vector is $\psi = \tan^{-1}(\frac{log(\lambda_1) \lambda_1^k}{log(\lambda_2) \lambda_2^k})+\frac{\pi}{2}$
\item for conjugate eigenvalue pairs: the pair forms a logarithmic spiral. The tangent of a spiral with polar equation $r=a e^{b\theta}$ will have a derivative $\frac{d r}{d \theta}=b a e^{b\theta}=b r$, hence the angle of the vector normal to the curve is $\tan^{-1}(\frac{r}{\frac{d r}{d \theta}})+\frac{\pi}{2}=\tan^{-1}(\frac{1}{b})+\frac{\pi}{2}$
\item for Jordan blocks with $gm>1$: in the case of real eigenvalues, we follow the same procedure as in the real case, but considering that the equation for each eigenvalue is $\frac{k^j}{j!}e^{\lambda k}$ where $j$ is the index of the term within the Jordan block (note that when $j=0$ the term is the same as for real eigenvalues above). The derivatives f'(k),g'(k) are in the form $(j+k*log(\lambda))\frac{k^{j-1}}{j!}e^{\lambda k}$.
\item for any combination of complex eigenvalues we use the above equations using the envelope of their magnitude (ie $|\lambda|$).
\end{enumerate}

@@

%  The na\"ive approach described in Sec.~\ref{sec:CEGIS-precision-incrementation}
  %uses a multi-staged verification phase and performs solution generalization based on
%the plant precision, the number of loop unrollings, 
  %and the sampling rate. Essentially, it
%  synthesizes a controller for an individual initial state
%and input with a bounded time horizon and, subsequently, it generalizes it to all reachable states,
%inputs, and time horizons during the verification phase.
Essentially, this approach relies on the symbolic
simulation over a bounded time horizon of individual initial states and inputs that form
part of an uncountable space and tries to generalize it for an
infinite space over an infinite time horizon.

@@

As stated in section \ref{ssec:stabspecification}, we use Jury's Criterion to assert this specification.

>>>>

\section{Introduction}

Modern implementations of embedded control systems have proliferated with
the availability of low-cost devices that can perform highly non-trivial
control tasks, with significant impact in numerous application areas such as
life sciences, environmental control and robotics~\cite{astrom1997computer,
Franklin15}.  Correct synthesis is non-trivial, even in cases with
restriceted dynamics.  We examine the case of Linear Time Invariant (LTI)
Models, for which the synthesis of controllers is well understood.  However,
the use of digital control architectures adds new challenges caused by
artifacts specific to digital control, such as the effects of
finite-precision arithmetic, time discretization and quantization noise
introduced by A/D and D/A conversion.

@@

\item We automatically generate {\em correct-by-construction} digital
  controllers using an inductive synthesis approach.  Our application of program
  synthesis is non-trivial and addresses challenges specific to control
  systems, such as quantizers and FWL.  In particular, we
  have found that a two-stage verification engine that continuously refines
  the precision of the fixed-point representation of the plant yields a
  speed-up of two orders of magnitude over a conventional one-stage
  verification engine.
%
\item  We compute state-feedback controllers that 
  guarantee a given safety property alongside stability and response time
  specifications.  Existing methods for controller synthesis rely on transfer
  function representations, which are inadequate to prove safety requirements.
%
\item We provide three novel algorithms: the first, na\"ive one, relies on
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction refinement and
  acceleration to improve scalability while retaining soundness.  A third
  algorithm evaluates systems using transfer functions where no state-space
  description is given.  All three approaches provide sound synthesis of
  feedback controllers and consider the various sources of imprecision in
  the implementation of the control algorithm and in the modeling of the
  plant.
%
\item We develop a model for different sources of quantization error and
  their effect on reachability properties.  We give bounds that ensure the
  safety of our controllers in a hybrid continuous-digital domain.
\end{enumerate}

@@

%-------------------------------
\subsection{Abstract Acceleration} 
\label{ssec:aa}
%-------------------------------

For abstraction based synthesis, we will use Abstract
Acceleration~\cite{JSS14,cattaruzza2015unbounded}.  Given an iterative
program with dynamics
%
\begin{equation}
\vec{x}_{k+1}=\mat{A}\vec{x}_k : \vec{x}_k \in \mathbb{R}^n \wedge \mat{A} \in \mathbb{R}^{n \times n} \wedge k \in [0\ \ \infty),
\end{equation}
%
we want to find all possible states visited over an infinite time horizon
starting at an initial set $X_0$.  We call this set the \emph{reach tube} of
the model:
%
\begin{equation}
\hat{X} = \{ \vec{x}_k : k \in [0\ \infty) \wedge \vec{x}_0 \in X_0 \wedge \vec{x}_{k+1}=\mat{A}\vec{x}_k\}.
\end{equation}
%
It is easy to see that the infinite number of iterations would result in an
unbounded cumulative rounding error, but using Abstract Acceleration we
transform this problem into a one-step solution.  We first change the
problem statement by means of acceleration into
%
\begin{equation}
\hat{X} = \{ \vec{x}_k=\mat{A}^k\vec{x}_0 : k \in [0\ \infty) \wedge \vec{x}_0 \in X_0 \}, 
\end{equation}
%
and further define new model semantics and corresponding set: 
%
\begin{equation}\label{eq:aa_reachtube}
\hat{X} \subseteq \hat{X}^\sharp = \mathcal{A}X_0, \text{ such that } \bigcup_{k \in [0\ \infty)} \mat{A}^k \subseteq \mathcal{A}, 
\end{equation}
%
where the newly introduced \emph{abstract reach tube} $\hat{X}^\sharp$ is an
over-approximation of the reach tube of the model \cite{JSS14}.  All we need
to do is define the nature of $\mathcal{A}X_0$.

@@

Introduce set $X=\{ \vec{x} : \mat{C}\vec{x} \leq \vec{d}\}$. 
The linear transformation $\vec{x}'=\mat{S}^{-1}\vec{x}$ leads to set 
%
\begin{equation}
X'= \mat{S}^{-1}X = \{\vec{x}' : \mat{C}\mat{S}\vec{x}'\leq \vec{d}\}.  
\label{eq:eigenInit}
\end{equation}
%
Analogously, let $X_0'=\mat{S}^{-1}X_0$ and
${\hat{X}'^\sharp}=\mat{S}^{-1}{\hat{X}^\sharp}$ be the initial set and the
abstract reach tube mapped onto the eigenspace of matrix $\mat{A}$.  We may
transform equation \eqref{eq:aa_reachtube} into:
%
\begin{equation}
\hat{X}' \subseteq {\hat{X}'^\sharp} = \mathcal{J}X_0', \text{ such that } \bigcup_{k \in [0\ \ \infty)}\mat{J}^k \subseteq \mathcal{J}, 
\end{equation}
%
where $\hat{X}'=\mat{S}^{-1}\hat{X}$ and $\mathcal{J}$ is a convex
polyhedron representing restrictions on the eigenvalues of $\mat{A}$. 
$\hat{X}'^\sharp$ is calculated by applying a simplex algorithm with a
tableau defined by $\mathcal{J}$ and objective functions derived from each
vertex of $X_0'$ (step \eqref{alg:getReachTube} in Algorithm
\ref{alg:findReachTube} below).  We note, however, that if we wish to use
numerical analysis in the decomposition of $\mat{A}$, the Tableau in
$\mathcal{J}$ contains small intervals given by the errors, which cannot be
processed efficiently by a regular simplex and require a new procedure.

The calculation of the abstract reach tube via abstract acceleration is
encompassed in Algorithm~\ref{alg:findReachTube}, which can be summarised as
follows (its steps are denoted in parentheses):
%
\begin{enumerate}
%
\item We perform unsound eigen-decomposition using an existing algebra package (lines \ref{alg:calcEigenvalues}-\ref{alg:calcEigenvectors}).
%
\item Then we restore soundness to the results using the methods described in Section \ref{sec:eigensolver} (lines \ref{alg:soundEigenvalues}-\ref{alg:soundEigenvectors}). 
%
\item The inverse of the matrix of eigenvectors is calculated after soundness is restored, using interval arithmetic in order to ensure its soundness (line \ref{alg:inverse}).
%
\item The abstract dynamics are obtained by evaluating the convex hull of all powers of eigenvalues up to the desired number of iterations as described in \cite{cattaruzza2015unbounded} (line \ref{alg:aa}).
%
\item Using Equation \eqref{eq:eigenInit}, transform the initial state into the eigenspace (line \ref{alg:eigeninit}).
%
\item Extract the vertices of the eigen-polyhedron $X_0'$ using the sound over-approximation of the double description algorithm \cite{fukuda1996double} from Section \ref{sec:dd} (line \ref{alg:getVertices}). It is worth noting that this algorithm uses a simplex to seed it: we will first discuss the simplex algorithm, which is also needed at the next step, and then the vertex enumeration algorithm used at this stage. 
%
\item Calculate the mapped \emph{abstract reach tube} ${\hat{X}'^\sharp}$
that over-approximates the image of the reach tube in the eigenspace (line
\ref{alg:getReachTube}).  This is achieved by evaluating a set of objective
functions using the abstract dynamics as a simplex Tableau, and via a sound
simplex described in Section \ref{sec:simplex}.  The objective functions are
defined by the vertices of $X'_0$, denote as $V_0$, and by the desired
template directions, such that
%
\begin{equation*}
\vec{w}_{ij}= \vec{v}_i \circ \vec{t}_j : \vec{v}_i \in V_0 \wedge \vec{t}_j \in T,
\end{equation*}
%
where $T$ is the set of template directions, and $\circ$ denotes a
component-wise multiplication yielding a vector.
%
\item Using Equation~\eqref{eq:eigenInit}, find the reach tube
$\hat{X}^\sharp=\mat{S}{\hat{X}'^\sharp}$ (line \ref{alg:getNormReachTube}).
%
\end{enumerate}

@@

%-------------------------------
\subsection{Overapproximating continuous time dynamics using discrete models}
\label{sec:cont_aa}
%-------------------------------

Thus far we wave discussed the Abstract Accelertion of Discrete Time
systems.  However, our plant models derive from the discretization of
continuous time systems.  We therefore seek to establish a sound
overapproximation which does not only encompass the selected discretization,
but any chosen discretization in the time domain.

A dynamical system is a system in which a function describes the progression
of a state over time.  In a continuous domain with linear dynamics, it is
described by a first order Ordinary Differential Equation (ODE).
%
\begin{equation}
\dot{x}(t)=\mat{A}_o\vec{x}(t)+\mat{B}_o\vec{u}(t) +\vec{w}(t)
\label{eq:dynamical}
\end{equation}
%
where $\vec{w}$  is a continuous zero-mean white noise source with
covariance $N ( 0 , Q )$ Furthermore, a control system may have a derived
output that is a linear combination of its states and inputs, which may
restricts the observability of the statespace from the output space.
%
\begin{equation}
\vec{y}(t)=\mat{C}_o\vec{x}(t)+\mat{D}_o\vec{u}(t)
\end{equation}
%
Discretization of a continuous dynamical system turns the ODE into a
difference equation, assuming zero-order hold for the input $\vec{u}$ and
continuous integration for the noise $\vec{w}$, to
%
\begin{align}
\label{eq:discretization}
\vec{x}_{k+1} &= \mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k + \vec{w}_k\\
y_k &= \mat{C}_d \vec{x}_ k + \mat{D}_d \vec{u}_ k 
\end{align}
with covariance for $\vec{w}_k$, $N ( 0 , Q_d )$
where
\begin{align}
\label{eq:discretize}
\mat{A}_d &= e^{\mat{A}_o T_s} = \mathcal{L}^{-1} { ( s \mat{I} - \mat{A}_o )^{-1} }_{t = T_s}\\
\mat{B}_d &= \int_{t = 0}^{T_s} e^{\mat{A}_o t} dt\ \mat{B}_o = \mat{A}_o^{-1} ( \mat{A}_d - \mat{I} ) \mat{B}_o\\
\mat{C}_d &= \mat{C}_o\\
\mat{D}_d &= \mat{D}_o \\
Q_d &= \int_{t = 0}^{T_s} e^{\mat{A}_o t} Q e^{\mat{A}_o^T t} dt
\end{align}
and $T_s$ is the sample time. Then
\begin{align*}
x(kT)=x_k \wedge y(kT) = y_k, \forall k
\end{align*}

@@

 Let $\mat{J}$ be a diagonal matrix, such that 
 $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
 Let us now take a different sample rate $T_2$, such that 
 $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
 We can then say that 
 \begin{equation}
 {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
 \end{equation}
 This proves the theorem for $gm(\lambda_i)=1$.
 We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
 referenced to the original eigenvalues modified by the sampling rate.
  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix. Then
\begin{align}
 \mat{J}_1=&\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k\nonumber\\
 =&\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
 \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
\vdots & \ddots & \ddots & \vdots \\
& &  &\lambda_i^k \\
\end{array} \right ] T_1^k
\end{align}
%
The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
%
\begin{align}
\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
\end{align}
which completes the proof for $gm(\lambda_i)>1$.
\end{proof}

@@

In the case of Discrete dynamics, support functions can be found by
evaluating hyperplanes of consecutive exponents of the eigenvalues using
differences.  In the case of continuous time, this translates to calculating
derivatives.  Once again, when the sets are known to be convex, we may
calculate the normal vector to this derivative and evaluate its support
function on the selected iteration.  The direction of the vector is
determined by evaluating the support function of another point in the curve,
which is known to be smaller than that of the selected point.  The
corresponding derivatives are:
%
\begin{enumerate}
%
\item for real eigenvalues: we first calculate the slope of the tangent
between the two progressions: $\frac{d f(k)}{d g(k)}=\frac{f'(k)}{g'(k)}$. 
Hence $\frac{d \lambda_1^k}{d \lambda_2^k}=\frac{log(\lambda_1)
\lambda_1^k}{log(\lambda_2) \lambda_2^k}$.  From this we know that the polar
angle of the support vector is $\psi = \tan^{-1}(\frac{log(\lambda_1)
\lambda_1^k}{log(\lambda_2) \lambda_2^k})+\frac{\pi}{2}$
%
\item for conjugate eigenvalue pairs: the pair forms a logarithmic spiral. 
The tangent of a spiral with polar equation $r=a e^{b\theta}$ will have a
derivative $\frac{d r}{d \theta}=b a e^{b\theta}=b r$, hence the angle of
the vector normal to the curve is $\tan^{-1}(\frac{r}{\frac{d r}{d
\theta}})+\frac{\pi}{2}=\tan^{-1}(\frac{1}{b})+\frac{\pi}{2}$
%
\item for Jordan blocks with $gm>1$: in the case of real eigenvalues, we
follow the same procedure as in the real case, but considering that the
equation for each eigenvalue is $\frac{k^j}{j!}e^{\lambda k}$ where $j$ is
the index of the term within the Jordan block (note that when $j=0$ the term
is the same as for real eigenvalues above).  The derivatives f'(k),g'(k) are
in the form $(j+k*log(\lambda))\frac{k^{j-1}}{j!}e^{\lambda k}$.
%
\item for any combination of complex eigenvalues we use the above equations
using the envelope of their magnitude (ie $|\lambda|$).
%
\end{enumerate}

@@

%  The na\"ive approach described in Sec.~\ref{sec:CEGIS-precision-incrementation}
  %uses a multi-staged verification phase and performs solution generalization based on
%the plant precision, the number of loop unrollings, 
  %and the sampling rate. Essentially, it
%  synthesizes a controller for an individual initial state
%and input with a bounded time horizon and, subsequently, it generalizes it to all reachable states,
%inputs, and time horizons during the verification phase.

Essentially, this approach relies on the symbolic simulation over a bounded
time horizon of individual initial states and inputs that form part of an
uncountable space and tries to generalize it for an infinite space over an
infinite time horizon.

@@

As stated in Section~\ref{ssec:stabspecification}, we use Jury's Criterion
to assert this specification.
